%%%% Background Section

\section{Background and Existing Work}
	% the robustness monitor approach in RV2014 is close to invariants
	% if you define your invariants well (and potentially add different levels of failures (e.g., warnings) you'd get to the same place
	% they require a predictive component in the system  -- we just want envelope
	% horizon is delay
	% observation map is semi-formal mapping
In this section we briefly introduce the background concepts and
discuss relevant existing work that will put the current work in perspective.

\vspace*{3pt}
\noindent
\textbf{Monitoring architecture.}
%\label{sec:bg:sc_monitor}
Goodloe and Pike present a thorough survey of monitoring distributed real-time systems
\cite{Goodloe2010}. Notably, they present a set of monitor architecture constraints and
propose three abstract monitor architectures in the context of monitoring these types of systems.
%
%In \cite{Pike2011} Pike et. al update these constraints with the acronym ``FaCTS'': Functionality, Certifiability, Timing, and SWaP (size, weight and power).
%These constraints emphasize the need for strong isolation between the target system and the monitor. Without isolation, the monitor may affect the target system in a way which could cause a system failure (e.g., disrupting system timing, adding extensive development costs, etc).
%The Functionality constraint demands that a monitor cannot change the system under observation's (SUO's) behavior unless the target has violated the system specification.
%The Timing constraint similarly says that the monitor can not interfere with the non-faulty SUO's timing (e.g., task period/deadlines).
%The Certifiability constraint is a softer constraint, arguing that a monitor should not make re-certification of SUO onerous. This is important because certification can be a major portion of design cost for these systems and nominally simple changes/additions to the SUO can require a broad and costly recertification.
%Lastly, safety critical systems are often extremely cost sensitive with tight tolerances for additional physical size, weight or required power. Any monitor we wish to add to an existing system must fit within these existing tolerances.
%The three monitor architectures proposed by Goodloe and Pike are the Bus-Monitor Architecture, the single process monitor architecture, and the distributed process monitor architecture.
One of the proposed distributed real-time system monitor architectures is the bus-monitor architecture.
This architecture contains an external monitor which receives network messages over an existing system bus, acting as another system component.
The monitor can be configured in a silent or receive only mode to ensure it does not perturb the system.
This is a simple architecture which requires minor changes to the target system.
We utilize this architecture for our monitoring framework.

%%%% NEED TO SHRINK THIS DOWN TO A PAGE OR TWO
%%%% 	focus on the actual similar algorithms (Thati/Rosu, Havelund, etc)

%%% Have to mention:
%%% Thati/Rosu [27] monitoring for mtl
%%% Basin MTL algorithms [5]
%%% Copilot
%%% MaC?? -- probably
%%% Reinbacher
%%% Heffernan
%%% Precis

\vspace*{3pt}
\noindent
\textbf{Controller area network.} Controller Area Network is a
widely used automotive network developed by Bosch
in the 1980s \cite{Bosch1991}. In this work we  primarily focus on CAN as
it is a common automotive bus which typically conveys enough of the state
information so that we can check for interesting safety properties of the system.
CAN is an event-based broadcast network with data rates up to 1Mb/s.
Messages on CAN are broadcast with an identifier which is used to denote
both the message and the intended recipients.
The message identifiers are also used as the message priorities for access control.
Although CAN is an event-based bus,
it is often used with periodic scheduling schemes so the network usage can be statically
analyzed. Hence, our monitoring approach is based on a time-triggered, network sampling
model which allows it to monitor time-triggered networks as well.
We add \monitor as a
passive external bus-monitor which can only check system properties
that are observable by passive
observation of the messages transmitted through CAN.




\vspace*{3pt}
\noindent
\textbf{Monitoring algorithm.}
Our monitoring algorithm is similar to existing dynamic programming and
formula-rewriting based algorithms \cite{Havelund2004,Havelund2002,Rosu2005,Thati2005,Basin2012}.
% need to get good novelty wording
% topics:
%	real-time eager checking of future properties
%	hybrid eager checking for efficiency
%	practical experience showing suitability of our language for safety monitoring
%The primary novelty in our approach is the combination of eager and delayed checking for real-time monitoring,
Our main area of novelty is the combination of eager and conservative specification checking used in a practical setting showing the suitability of our bounded future logic for safety monitoring.
%\textit{Dynamic programming monitors.}
%% garg/precis
Our monitoring algorithm is inspired by the algorithms \greduce \cite{Garg2011} and
\precis \cite{Chowdhury2014}, adjusted for propositional logic and eager checking.
The structure of our algorithm is based on \greduce.
% We utilize an iterative, formula-rewriting based algorithm targeted at both offline log analysis as well as runtime monitoring.
\greduce, \precis, and \monitor can handle future incompleteness but \greduce additionally
considers incompleteness for missing information which we do not consider.
% \precis and \monitor both require the input trace to contain complete information.
%
% return residual (i.e., incompletely reduced) formulas, but incompleteness in \greduce\ is due to incomplete logs which lead to unknown predicate substitutions, whereas our algorithm works on complete logs but must deal with temporal (i.e., future-time) incompleteness.
%
% % dynamic prog algos -- thati/rosu, havelund, etc
The NASA PathExplorer project has led to both a set of dynamic programming-based monitoring algorithms as
well as some formula-rewriting based algorithms \cite{Havelund2004} for past-time LTL.
% These dynamic programming algorithms require checking the trace in reverse
% (from the end to the beginning) which makes them somewhat unsuitable for online monitoring \cite{Havelund2002}.
The formula rewriting algorithms utilize the Maude term rewriting engine to efficiently monitor specifications
through formula rewriting \cite{Rosu2005}.
%
Thati and Ro\c{s}u \cite{Thati2005} describe a dynamic programming and rewriting-based
algorithm for monitoring MTL formulas.
% which is based on resolving the past and deriving the future.
They perform eager runtime monitoring by formula rewriting which resolves past-time formulas
into equivalent formulas without unguarded past-time operators and derive new future-time
formulas which separate the current state from future state.
While they have a tight encoding of their canonical formulas, they still require more memory
than some existing algorithms (formulas
grow in size as they are rewritten), including
\monitor.
% the current work.
%Their evaluation algorithm is similar to ours, but they require more storage space to handle formula rewriting with potentially growing formulas.


%\textit{Embedded Monitors.}
%% Hef/Rein
Heffernan et. al. present a monitor for automotive systems using ISO 26262 as a guide
to identify the monitored properties  \cite{Heffernan2014}.
They monitor past-time linear temporal logic (LTL) formulas and obtain system state from target system buses (CAN in their example).
Our \sfmap component is similar to their ``filters'' used to translate system state to the atomic propositions used in the policy.
Their motivation and goals are similar to ours, but they use system-on-a-chip based
monitors which utilize instrumentation to obtain system state, which is not suitable for monitoring black-box commercial-off-the-shelf (COTS)
systems. Reinbacher et. al. present an embedded past-time MTL monitor in \cite{Reinbacher2013} which generates FPGA-based non-invasive monitors.
The actual implementation they describe does however presume system memory access to obtain system state (rather than using state from the target network).
Pellizzoni et. al. describe a monitor for COTS peripherals in \cite{Pellizzoni2008}.
They generate FPGA monitors that passively observe PCI-E buses to verify system properties.
Although their architecture is similar to ours, but they only check past-time LTL and
regular expressions which cannot capture timing properties.
%%%%%%%%%%%%%%
Basin et. al. compare runtime monitoring algorithms for MTL properties \cite{Basin2012}.
\monitor works similarly to their point-based monitoring algorithm but \monitor checks
future temporal operators more aggresively.
Donz\'e \etal \cite{DFM13} developed a robustness monitor for Signal Temporal Logic (STL) which
supports continuous signals.
Nickovic and Maler \cite{NM07} developed the  AWT tool which monitors analog systems.
We only consider discrete events. Dokhanchi \etal \cite{DHF14}
developed an online runtime monitoring algorithm for checking the robustness of formulas written in a
future-bounded MTL fragment. We use the same MTL fragment but we consider satisfaction of the formula
instead of robustness.
% Though they discuss the use of
% delay queues to monitor future-time properties (and thus do not eagerly check future-time formulas), we could integrate their algorithm into our eager monitoring framework.
