\documentclass{article}

\title{LTLFonts}
\author{Matteo Slanina}
\date{January 2007}

\usepackage{ltlfonts,booktabs,url}

\begin{document}

\maketitle


\section{License}

LTLFonts is released under the \emph{SIL Open Font License}; please
read it carefully and do not use the fonts unless you agree to the
terms of the license:

This Font Software is Copyright (c) 2006--2007, Matteo Slanina
(\url{http://theory.stanford.edu/~matteo/ltlfonts/}). All Rights
Reserved.

``LTLFonts'' is a Reserved Font Name for this Font Software.

This Font Software is licensed under the SIL Open Font License,
Version 1.0. No modification of the license is permitted, only
verbatim copy is allowed. This license is included in the LTLFonts
distribution, and is also available with a FAQ at:
\url{http://scripts.sil.org/OFL}


\section{Usage}

It's as easy as
\begin{verbatim}
\usepackage{ltlfonts}
\end{verbatim}

Then run your source through \LaTeX{} as usually, and call
\verb|dvips| with the option \verb|-u +ltlfonts.map|, or add the line
\begin{verbatim}
p +ltlfonts.map
\end{verbatim}
to your \verb|.dvipsrc|.

If you use a different back-end than \verb|dvips|, you must probably
create a similar map file.


\section{Glyph Complement}

All commands are math-mode only. They are long and inconvenient on
purpose: you are supposed to define semantically-oriented commands
according to your own preferences.

\subsection*{``Regular'' Symbols}

The following symbols are the ones you are most likely to use. They
match well with lower-case math letters:
\begin{center}
  \begin{tabular}{@{}cll@{}}
    \toprule
    glyph & command \\
    \midrule
    $\mathstrut\LTLsquare         $ & \verb|\LTLsquare         | \\
    $\mathstrut\LTLdiamond        $ & \verb|\LTLdiamond        | \\
    $\mathstrut\LTLcircle         $ & \verb|\LTLcircle         | \\
    $\mathstrut\LTLsquareminus    $ & \verb|\LTLsquareminus    | \\
    $\mathstrut\LTLdiamondminus   $ & \verb|\LTLdiamondminus   | \\
    $\mathstrut\LTLcircleminus    $ & \verb|\LTLcircleminus    | \\
    $\mathstrut\LTLcircletilde    $ & \verb|\LTLcircletilde    | \\
    $\mathstrut\LTLsquarehat      $ & \verb|\LTLsquarehat      | \\
    $\mathstrut\LTLdiamondhat     $ & \verb|\LTLdiamondhat     | \\
    $\mathstrut\LTLsquareminushat $ & \verb|\LTLsquareminushat | \\
    $\mathstrut\LTLdiamondminushat$ & \verb|\LTLdiamondminushat| \\
    \bottomrule
  \end{tabular}
\end{center}
For example, \verb|\LTLsquare \phi| produces
\[ \LTLsquare \phi \]
and \verb|\LTLdiamondminus \LTLsquareminus (x = 0)| produces
\[ \LTLdiamondminus \LTLsquareminus (x = 0) \]

\subsection*{``Uppercase'' Symbols}

The following symbols are identical to the previous ones, but raised
slightly to make them align better with uppercase math letters. You
can use them in specific places to improve the presentation, if you
are really picky.

Please notice that LTLFonts are not designed to be used in an
``upper-case'' prevalent notation, since neither case will align well
with operators on the math line such as ``$\rightarrow$.''
\begin{center}
  \begin{tabular}{@{}cl@{}}
    \toprule
    glyph & command \\
    \midrule
    $\mathstrut\LTLsquareuc         $ & \verb|\LTLsquareuc         | \\
    $\mathstrut\LTLdiamonduc        $ & \verb|\LTLdiamonduc        | \\
    $\mathstrut\LTLcircleuc         $ & \verb|\LTLcircleuc         | \\
    $\mathstrut\LTLsquareminusuc    $ & \verb|\LTLsquareminusuc    | \\
    $\mathstrut\LTLdiamondminusuc   $ & \verb|\LTLdiamondminusuc   | \\
    $\mathstrut\LTLcircleminusuc    $ & \verb|\LTLcircleminusuc    | \\
    $\mathstrut\LTLcircletildeuc    $ & \verb|\LTLcircletildeuc    | \\
    $\mathstrut\LTLsquarehatuc      $ & \verb|\LTLsquarehatuc      | \\
    $\mathstrut\LTLdiamondhatuc     $ & \verb|\LTLdiamondhatuc     | \\
    $\mathstrut\LTLsquareminushatuc $ & \verb|\LTLsquareminushatuc | \\
    $\mathstrut\LTLdiamondminushatuc$ & \verb|\LTLdiamondminushatuc| \\
    \bottomrule
  \end{tabular}
\end{center}
For example, \verb|\LTLsquareuc \Phi| produces
\[ \LTLsquareuc \Phi \]
(compare with \verb|\LTLsquare \Phi| = $\LTLsquare \Phi$).

\subsection*{Hat Accent}

\verb|\LTLhat| provides a math accent identical to the one on top of
\verb|\LTLsquarehat| and company, so you can build matching binary LTL
symbols. It comes in one width only, and will look good on top of
cursive ``U,'' ``W,'' ``S,'' and ``B,'' but may be off-center on top
of other letters, as it wasn't really designed as a generic math
accent:
\begin{center}
  \begin{tabular}{@{}cl@{}}
    \toprule
    glyph & command \\
    \midrule
    $\mathstrut\LTLhat{\mathcal{U}}$ & \verb|\LTLhat{\mathcal{U}}| \\
    \bottomrule
  \end{tabular}
\end{center}
For example, \verb|p \mathbin{\LTLhat{\mathcal{B}}} q| produces
\[ p \mathbin{\LTLhat{\mathcal{B}}} q \]


\section{Kerning}

The unary operators are declared in \verb|ltlfonts.sty| to be of math
class 0 (``Ordinary'', \verb|\mathord|). This allows \TeX{} to use
kerning information: a few pairs, such as $\LTLdiamond\LTLcircle$, are
kerned to make the spacing look more uniform. Be aware that, if you
redeclare the symbols to be of another type, \TeX{} will not be able
to use the kerning tables (if you have to, don't worry, the
differences are minimal). The suggested command definition in your
document is thus something like
\begin{verbatim}
\newcommand\Henceforth{\LTLsquare}
\end{verbatim}


\section{Notes}

\begin{itemize}
\item Characters with hats extend above the font's declared ascender
  length. You should thus not set them solid (with
  \verb|\baselineskip| equal to the type size) or smaller in text
  longer than a single line.
\end{itemize}

\end{document}
